---
title: "Netflix Prices Around the World"
author: "Cliff Anderson"
date: "Spring 2022"
output: html_document
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      output_file = 'index.html') })
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
require(plotly)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


##### Background
Sherlock Holmes series is taken out from Netflix USA, but still appears in Netflix from Indonesia. Because of this, I became curious on how many TV series/movies are on Netflix for different countries. In addition, I also realize that prices are different, depending on where you are. 

Source: https://www.kaggle.com/prasertk/netflix-subscription-price-in-different-countries 

```{r, echo = FALSE}
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')
colnames(df) <- c('Country')
df2 <- read.csv('netflix_price.csv')

df3 <- merge(df, df2)

# Rename columns
names(df3)[3] <- 'Code'
names(df3)[7] <- 'Cost_basic'
names(df3)[8] <- 'Cost_standard'
names(df3)[9] <- 'Cost_premium'
df3 <- df3[ -c(2) ]

# Add column cost per title (Basic plan)
df3 <- df3 %>%
  select(Country,Code,Total.Library.Size,No..of.TV.Shows,No..of.Movies,Cost_basic,Cost_standard,Cost_premium) %>%
  mutate(Cost_per_title_basic = Cost_basic / Total.Library.Size)



```


```{r, echo = FALSE}
fig <- plot_ly(df3, type='choropleth', locations=df3$Code, z=df3$Cost_basic, text=df3$Country, colorscale="Blues")

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)


fig <- fig %>% colorbar(title = 'Cost (US$)', tickprefix = '$')
fig <- fig %>% layout(
    title = 'Netflix Monthly Basic Subscription<br> Source: <a href="https://www.kaggle.com/prasertk/netflix-subscription-price-in-different-countries">Kaggle</a>'
  )
fig
```



```{r, echo = FALSE}
# scatter plot comparing price and library size
fig2 <- plot_ly(data = df3, x = ~Cost_basic, y = ~Total.Library.Size,
               # Hover text:
               text = ~paste("Price: $", Cost_basic, '<br>Library Size:', Total.Library.Size, "<br>Country:", Country),
               marker = list(size = 10,
                             color = 'rgba(255, 182, 193, .9)',
                             line = list(color = 'rgba(152, 0, 0, .8)',
                                         width = 2)))

fig2 <- fig2 %>% layout(title = 'Total Library Size vs Cost',
         yaxis = list(zeroline = FALSE, title = 'Total Library Size (Movie + TV Show)'),
         xaxis = list(zeroline = FALSE, title = 'Monthly Subscription Cost $'))

fig2
```


```{r, echo = FALSE, fig.width = 13, fig.height = 7}
bar_fig <- plot_ly(
  data = df3,
  y = ~Cost_per_title_basic,
  x = ~Country,
  type = 'bar'
) %>% 
  layout(title = 'Cost per title for the Basic Plan',
         xaxis = list(categoryorder = "total descending"),
         yaxis = list(title = 'Cost in USD'), barmode = 'group'
         )
bar_fig
```


##### Reflection
In Chapter 15.3, Wilke suggested that for online viewing, dark-colored background areas may be preferable. This means that the higher the variable is, the lighter the color. I implemented this on my Chloropleth map. 
One thing I wish I can do is have a filter bar. For example, filter countries whose price is less than $5. I also wish I can categorize the countries into groups of continents so that I can analyze the data that way. 

#### Part 2
Added a bar graph to show how much one movie/TV series cost for the basic monthly plan. Since countries have different Netflix cost and different number of movies, I wanted to see which country has the "best deal". 
  