---
title: "I Speak for the Trees"
author: "Samantha Driesenga sld39"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
library(plotly)
library(tidyverse)
```


### Topic Background

Because I am a big environmentalist, I chose the topic of deforestation. To do this, I chose to
plot the change in forest cover (in percentage) for each individual continent from 1990-2020.

### Data Overview

The data I used for this visual was originally published by the UN Food and Agriculture
Organization (FAO).

Original published form: https://fra-data.fao.org/WO/assessment/fra2020 

Where I obtained the data:
https://ourworldindata.org/grapher/forest-area-as-share-of-land-area?tab=chart&time=earliest..latest&country=England~Scotland~FRA~USA


```{r load-data, include=FALSE}
forest_data <- readr::read_csv('forest-area-as-share-of-land-area.csv')
summary(forest_data)
nrow(forest_data)
#forest_data
glimpse(forest_data)
```

```{r group, include=FALSE}
forest_data %>%
  group_by(Entity)
```
```{r ungroup, include=FALSE}
forest_data %>%
  ungroup
```


```{r wrangle, include=FALSE}
wrangled_forest <- forest_data %>%
  mutate(land_type = as.factor(ifelse(Entity == 'World', 'World',
                            ifelse(!is.na(Code), 'Country', 
                            ifelse(Entity %in% c('Africa' , 'Asia' , 'Central America' , 'Europe' , 'Northern America',  'Oceania' , 'South America') , 'Continent' , 
                                   'Region')))))
wrangled_forest %>%
  arrange(desc(Entity))
```
```{r, include=FALSE}
wrangled_forest <- wrangled_forest %>%
  filter(land_type == 'Continent')
wrangled_forest
```

```{r, include=FALSE}
Africa_forest <- wrangled_forest %>%
  filter(Entity == 'Africa')
```

```{r, include=FALSE}
Asia_forest <- wrangled_forest %>%
  filter(Entity == 'Asia')
```

```{r, include=FALSE}
CA_forest <- wrangled_forest %>%
  filter(Entity == 'Central America')
```

```{r, include=FALSE}
Europe_forest <- wrangled_forest %>%
  filter(Entity == 'Europe')
```

```{r, include=FALSE}
NA_forest <- wrangled_forest %>%
  filter(Entity == 'Northern America')
```

```{r, include=FALSE}
Oceania_forest <- wrangled_forest %>%
  filter(Entity == 'Oceania')
```

```{r, include=FALSE}
SA_forest <- wrangled_forest %>%
  filter(Entity == 'South America')
```

```{r, include=FALSE}
continent_forest <- data.frame(Africa_forest, Asia_forest, CA_forest, Europe_forest, NA_forest, Oceania_forest, SA_forest)
```

```{r, include=FALSE}
continent_forest <- continent_forest %>%
  group_by('Entity')
```

```{r, echo=FALSE}
plot_vis <- plot_ly(continent_forest, x = ~Year, y = ~forest_cover, name = 'Africa', type = 'scatter', mode = 'lines')

plot_vis <- plot_vis %>% add_trace(y = ~forest_cover.1, name = 'Asia', type = 'scatter', mode = 'lines')
plot_vis <- plot_vis %>% add_trace(y = ~forest_cover.2, name = 'Central America', type = 'scatter', mode = 'lines')
plot_vis <- plot_vis %>% add_trace(y = ~forest_cover.3, name = 'Europe', type = 'scatter', mode = 'lines')
plot_vis <- plot_vis %>% add_trace(y = ~forest_cover.4, name = 'North America', type = 'scatter', mode = 'lines')
plot_vis <- plot_vis %>% add_trace(y = ~forest_cover.5, name = 'Ociania', type = 'scatter', mode = 'lines')
plot_vis <- plot_vis %>% add_trace(y = ~forest_cover.6, name = 'South America', type = 'scatter', mode = 'lines')

plot_vis <- plot_vis %>% layout(title = "Continental Forest Coverage",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Forest Coverage (Percentage)"))



plot_vis
```

### Reflection

What I took from Wilke to use in this project is how we are telling a story. Though fighting
deforestation is a global effort, it isn't necessarily globally caused. There are places in 
the world in a state of reforestation, though deforestation does continue to outweigh
reforestation.
I wish I was able to show the change in forest cover numerically as well as graphically.
Perhaps in my next attempt I'll be able to include the change from the previous year in my
traces.
What I found most frustrating was the amount of coding required to make the graph. I made
nearly an exactly identical graph of this data for a previous project in about half the amount
of code.
