---
title: "Red Project Grand Rapids - plotly Project"
author: "Sonali Bola" 
date: "February 16th, 2022" 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic) 
```

## Data: 
```{r}
rp_raw <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTC3DmH_qt8vUg_WuWgSlfskPljYA-g_wlOfmpoXHJGUDIx_k2NHO4rwPh28QffW_govFRjAxFVjFxp/pub?gid=0&single=true&output=csv")

rp <- rp_raw %>%
  select(-"INCIDENT.NUMBER", -"NAME", -"INCIDENT.ADDRESS",  -"PHONE.NUMBER", -"RESIDENT.ADDRESS", -"EMESIS", -"COMBATIVE", -"DISCOMFORT", -"OTHER", -"ADDITIONAL.NOTES.FROM.REPORTS", -"SUD.SERVICES", -"MAT.SERVICES", -"HEALTH.INSURANCE", -"GENERAL.MEDICAL.SERVICES", -"HOUSING.RESOURCES", 
- "FOOD.RESOURCES", -"TRANSPORTATION.RESOURCES", -"OCCUPATIONAL.SERVICES", -"GENERAL.FOLLOW.UP.NOTES")

rp <- rp%>%
  filter(GENDER %in% c('M', 'F', 'UNK'))
```

a. Your visualization(s). Your visualization(s) should include at least one interactive element and follow good practices for the display of data.
b. Any background information needed to understand the data or the visualization.
c. A description of the data source with links if the data are available online. (This would include a link to data in this repository if that where the data live.)
d. Any other references/citations as needed.
e. A section reflecting on your project.
- What ideas/suggestions from Claus Wilke’s helped shape your visualization?
- Is there anything more you wish you could do with this data?
- What were the most interesting or frustrating technical aspects of doing this?
